---
title: "Whole Foods"
author: "Chunzi Wang"
date: "March 17, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(treemap)
```

```{r}
transaction <- read.csv("Whole_Foods_Transaction_Data.csv")
```

```{r}
head(transaction)
str(transaction)
```

```{r}
summary(transaction)
```

There're 24896 unique transactions. 

```{r}
length(unique(transaction$transaction_id))
```

```{r}
transaction %>%
  group_by(transaction_id) %>%
  count()

#same
#transaction %>%
  #count(transaction_id)
```

Find bestsellers:

```{r}
bestsellers <- transaction %>%
  group_by(product_name) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
bestsellers
```

23323 kinds of products in dataset. Groceries (fruits and vegetables) are the top 20.

```{r}
bestsellers %>%
  top_n(20) %>%
  ggplot(aes(x=reorder(product_name,-count),y=count,fill=count))+
  geom_col(show.legend = FALSE)+
  theme(axis.text.x=element_text(angle=90,hjust=1),axis.title.x=element_blank())
```

Find pattern in reordered items:

```{r}
transaction %>%
  group_by(reordered) %>%
  summarize(count=n())
```

```{r}
reordered <- transaction %>%
  filter(reordered==1) %>%
  group_by(product_name) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
reordered
```

Large overlap of reordered items and bestsellers. Check the propability of each item being reordered.

Set the count threshold to 100 since we're more interested in products that are more popular with the general public. This time, milk especially, and water, eggs, soda, coconut water, yogurt are the highest propability of reordered. Consider it beverage category. People who drink milk tend to drink it every day and refill it constantly.

```{r}
proportion_reordered <- transaction %>%
  group_by(product_name) %>%
  summarise(proportion_reordered = mean(reordered), count=n()) %>%
  filter(count>100) %>%
  arrange(desc(proportion_reordered))
proportion_reordered
```

There're more than 4000 kinds of products that're 100% reordered but the total times they are bought are comparatively small. They serve niche market but the customers who bought them are extremely loyal.

```{r}
transaction %>%
  group_by(product_name) %>%
  summarise(proportion_reordered = mean(reordered), count=n()) %>%
  filter(proportion_reordered==1) %>%
  arrange(desc(count))
```

Association between number of orders and probability of reordering, although some small countes of orders have high reordered rate, in general, there's a upward trend that shows a high number of orders are naturally more likely ro be reordered.

```{r}
proportion_reordered %>%
  ggplot(aes(x=count,y=proportion_reordered))+
  geom_point()+
  geom_smooth(color="red")
```

Visualizing product porfolio

```{r}
portfolio <- transaction %>%
  group_by(department,aisle) %>%
  summarise(count=n()) %>%
  ungroup() %>%
  group_by(department) %>%
  mutate(dep_sumcount=sum(count)) %>%
  ungroup() %>%
  mutate(onesize=1)
portfolio
```

```{r}
treemap(portfolio,index=c("department","aisle"),vSize="onesize",vColor="department",palette="Set3",title="",sortID="-dep_sumcount",border.col="#FFFFFF",type="categorical",fontsize.legend=0,bg.labels="#FFFFFF")
```

How oftem are products from each department/aisle sold?

It's clear that produce and diary eggs are the top 2 departments of all the transactions.

```{r}
treemap(portfolio,index=c("department","aisle"),vSize="count",title="",palette="Set3",border.col="#FFFFFF")
```

```{r}
portfolio %>%
  arrange(desc(count)) %>%
  top_n(20,count) %>%
  ggplot(aes(x=reorder(aisle,-count),y=count,fill=department))+
  geom_col()+
  theme(axis.text.x=element_text(angle=90,hjust=1),axis.title.x=element_blank())

```

Association Rules

```{r}
library(arules)
library(arulesViz)
```

```{r}
baskets <- transaction %>%
  group_by(transaction_id) %>%
  summarise(basket=as.vector(list(product_name)))
```

```{r}
class(baskets$basket)
length(baskets$basket)
```

```{r}
baskets$basket[[1]]
```

```{r}
transactions <- as(baskets$basket,"transactions")
```

Look at frequent items:

```{r}
item_frequencies <- itemFrequency(transactions,type="a")
support <- 0.02
freq_items <- sort(item_frequencies,decreasing = F)
freq_items <- freq_items[freq_items>support*length(transactions)]
length(freq_items)
```

```{r}
par(mar=c(2,10,2,2)); options(scipen=5)
barplot(freq_items,horiz = T, las=1, main = "Frequent Items",cex.names = 0.5)
abline(v=support*length(transactions),col="red")
```

Frequent itemsets:

```{r}
support <- 0.008
itemsets <- apriori(transactions, parameter=list(target="frequent itemsets",supp=support,minlen=2),control=list(verbose=FALSE))
```

```{r}
par(mar=c(5,18,2,2)+0.1)
sets_order_supp <- DATAFRAME(sort(itemsets,by="support",decreasing = F))
barplot(sets_order_supp$support,names.arg = sets_order_supp$items,xlim=c(0,0.02),horiz=T,las=2,cex.names = 0.8,main="Frequent Itemsets")
```

Use a low support threshold and a high confidence to generate strong rules even for items that are less frequent.

```{r}
rules1 <- apriori(transactions,parameter=list(support=0.0001,conf=0.6,maxlen=3),control=list(verbose=FALSE))
summary(rules1)
```

```{r}
plot(rules1)
```

There're some rules with a heavy lift indicating a strong association between the items. Further investigate those critical rules:

```{r}
inspect(sort(rules1,by="lift")[1:10])
```

Different flavors of the same type of food have strong association between each other. Coffee creamer and pasta sauce has strong association towards chicken.

```{r}
inspect(sort(rules1,by="confidence")[1:10])
```


Now we increase the support and decrease condidence to get rules of some more frequent items but with less confidence.

```{r}
rules2 <- apriori(transactions,parameter=list(support=0.001,conf=0.4,maxlen=3),control=list(verbose=FALSE))
summary(rules2)
```

```{r}
plot(rules2)
```

```{r}
inspect(sort(rules2,by="lift")[1:10])
```

Same type of items have high positive correlation. 

```{r}
inspect(sort(rules2,by="confidence")[1:10])
```

Yogurt, fruits, and drinks are often bought together.

Finally, further increase support and decrease confidence.

```{r}
rules3 <- apriori(transactions,parameter=list(support=0.005,conf=0.1,maxlen=3),control=list(verbose=FALSE))
summary(rules3)
```

```{r}
plot(rules3)
```

```{r}
inspect(sort(rules3,by="lift")[1:10])
```

Fruits and vegetables are highly correlated.

```{r}
inspect(sort(rules3,by="confidence")[1:10])
```

Lots of fruits. Banana is the most popular fruit among all. 




